---
author: "Zahlen Zbinden"
title: "Lab 1"
format: "pdf"
date: "04-06-2024"
---

```{r}
library(ggplot2)
library(tidyverse)
```


```{r}
mpg |>
  ggplot(aes(x=displ, y=hwy)) +
    geom_point()
```


```{r}
ggplot(data=mpg)
```

```{r}
str(mpg)
```

1. A blank graph
2. 234 x 11
3. Categorical, type of drive train. front-wheel, rear-wheel, 4wd
4. find below
5. It just shows all the unique pairs of class and drive. It doesn't have any numerical values that would lead to a deeper understanding of the data.


```{r}
mpg |>
  ggplot() +
    geom_point(aes(x=cyl, y=hwy))
```


```{r}
mpg |>
  ggplot() +
    geom_point(aes(x=drv, y=class))
```


```{r}
mpg |>
  ggplot() +
    geom_point(aes(x=displ, y=hwy, color=class, size=cty))
```

```{r}
mpg |>
  ggplot() +
    geom_point(aes(x=displ, y=hwy, size=class))
```

```{r}
mpg |>
  ggplot() +
    geom_point(aes(x=displ, y=hwy, color=class, shape=class))
```

Notice that aes is a mapping, and mapping are displayed in the legend


```{r}
mpg |>
  ggplot() +
    geom_point(aes(x=displ, y=hwy, color="blue"))
```

opposed to


```{r}
mpg |>
  ggplot() +
    geom_point(aes(x=displ, y=hwy), color="blue")
```

```{r}
str(mpg)
```

```{r}
sapply(mpg, function(x) length(unique(x)))
```

1. the points are not blue because we mapped color to the constant "blue", we didn't set a color outside of the aes call.
2. manufacturer, model, cyl, trans, drv, fl, class are categorical variables
3. 
  - color to categorical or to continuous gets a gradient
  - size to categorical will give a different size for each category
  - size to continuous (displ) will bin the values and give a different size for each bin
  - shape to categorical will give a different shape for each category
  - shape to continuous is not recomended as shapes are discrete and continuous is not, also there is a limited number of shapes that can be displayed
4. You get a perfect 45 degree line because you are plotting the same variable against itself.
5. This aes is easier to see if you set the shape to something like a circle with no fill, stroke controls the size of the border of the circle.
6. You get different colors based on the logical that you provided.


```{r}
mpg |>
  ggplot() +
    geom_point(aes(x=displ, y=displ, color=displ < 5), shape=21)
```

