---
title: "Homework 2"
author: Zahlen Zbinden
format: pdf
---

```{r setup}
#| message: false
library(tidyverse)
library(RColorBrewer)
```

Tasks that require an answer are bolded (inside `**` in the .Rmd file). For any task that includes a question (i.e. it ends with "?"), you should also answer the question in sentence form.

# Looking at Data

The following tasks all relate to the dataset `starwars` that comes with the dplyr package. Since, dplyr is included in the tidyverse, you don't need to import this data, just type its name to see it:

```{r}
starwars
```

It contains attributes of the characters in *some* of the Star Wars films. You can find out more about the variables by examining the help page:

```{r}
#| eval: false
?starwars
```

## 1.

**How many rows and columns does the dataset `starwars` have?** *(Use the output from the above chunk to answer this question)* (1pt)

The starwars data set has 87 rows, and 14 attributes (columns)

## 2.

The following code extracts the characters from the starwars data are not human and saves the result to a variable called `not_humans`.

```{r}
not_humans <- filter(starwars, species != "Human")
```

**How many non-human characters are in the dataset?** (2pts) (*Your answer should include code, relevant output, and a complete sentence answer to the question.*)

The not_humans subset of the starwars data set contains 48 non_human observations.

```{r}
nrow(not_humans)
```

## 3.

**Create a scatterplot of all (human and non-human) characters' mass against their height, using color to represent sex.** (2pts)

```{r}
ggplot(starwars, aes(x = mass, y = height, color = sex)) +
    geom_point(size = 2, alpha = 0.5, na.rm = T) +
    labs(title = "Mass Vs. Height", x = "Mass", y = "height") +
    scale_color_brewer(palette = "Set1") +
    theme(
        text = element_text(size = 15),
        axis.title.y = element_text(angle  = 0, vjust  = 0.5)
    )
```

## 4.

**Make a scatter plot of the non-human characters mass against their height.** (1pt)

```{r}
ggplot(not_humans, aes(x = mass, y = height)) +
    geom_point(size = 2, alpha = 0.5, na.rm = T) +
    labs(title = "Non-Human Mass Vs. height", x = "Mass", y = "Height") +
    scale_color_brewer(palette = "Set1") +
    theme(
        text = element_text(size = 15),
        axis.title.y = element_text(angle = 0, vjust = 0.5)
    )
```

## 5.

**Make a barchart of all (human and non-human) characters' eye colors.** (1pt)

```{r}
custom_colors <- c(
    "black" = "black", 
    "blue" = "blue", 
    "blue-gray" = "darkslategray4", 
    "brown" = "chocolate4", 
    "dark" = "bisque4",
    "gold" = "darkgoldenrod1",
    "green, yellow" = "darkolivegreen2",
    "hazel" = "#C9C789",
    "orange" = "darkorange1",
    "pink" = "deeppink",
    "red" = "red",
    "red, blue" = "darkorchid",
    "unknown" = "azure4",
    "white" = "azure",
    "yellow" = "yellow"
)
ggplot(starwars, aes(x = eye_color, fill = eye_color)) +
    geom_bar() +
    scale_fill_manual(values = custom_colors) +
    labs(title = "Count of Eye Colors", x = "Eye Colors", y = "Count") +
    guides(fill = guide_legend(ncol = 2)) +
    theme(
        text = element_text(size = 10),
        axis.title.y = element_text(angle = 0, vjust = 0.5),
        axis.text.x = element_text(angle = 45)
    )
```

## 6.

**Make a histogram of all (human and non-human) character's mass.** (1pt)

```{r}
ggplot(starwars, aes(x = mass)) +  
    geom_histogram(bins = 50) +
    labs(title = "Characters Mass", x = "Mass", y = "Count") +
    theme(
        text = element_text(size = 15),
        axis.title.y = element_text(angle = 0, vjust = 0.5)
    )
```

------------------------------------------------------------------------

(2pts) For correct `author` in header, and submitting both PDF and Quarto files.
