---
title: "Homework 1"
author: Zahlen Zbinden
format: pdf
---

```{r setup}
#| message: false
library(tidyverse)
```

Tasks that require an answer are bolded (inside `**` in the .qmd file). For any task that includes a question (i.e. it ends with "?"), you should also answer the question in sentence form.

# Quarto

## 0.

Before making any changes to this document, **Render** it to verify you have all the necessary components installed. You can double check the PDF output you get against the `homework_01.pdf` provided on the Assignment page in canvas.

If you run into any difficulties with this first step, ask for help ASAP.

## 1.

**Edit the `author:` line in the top section of this document to replace my name with yours**. (1 pt)

It's good practice to check changes as you go. So, once you've made the change, **Render** this document again, and check the document renders without error, and your name has indeed replaced mine in the PDF output.

# Reading and Writing R Code

## 2.

**Write code that accomplishes the following steps in this order:** (4 pts)

1.  Assign a variable called `temp_f` the value `32`
2.  Subtract 32 from `temp_f`, then divide it by 1.8. Assign the result to a variable called `temp_c`
3.  Call the function `paste()` with `temp_f` as the first argument, `temp_c` as the second argument and the `sep` argument set to `"F converted to C is "`.

```{r}
# Your code goes inside this code chunk
temp_f <- 32
temp_c <- (temp_f - 32) / 1.8
paste(temp_f, temp_c, sep = "F converted to C is ")
```

If your code is correct, the result of running the chunk should be:

    [1] "32F converted to C is 0"

## 3.

This week, you've seen how naming the arguments in a call to a function is optional, but can increase the readability of the code. For instance, this line of code doesn't name any of the arguments to `log()`:

```{r}
#| results: hide
log(1, 2)
```

Whereas, this code, which does the exact same thing, names all of its arguments explicitly:

```{r}
#| results: hide
log(x = 1, base = 2)
```

**Re-write each line of code in the chunk below to explicitly name all of the arguments:** (4 pts)

```{r}
#| fig-keep: none
#| results: hide
choose(n = 4, k = 2)

mean(x = LakeHuron, trim = 0.1, na.rm = TRUE)

ggplot(data = diamonds, mapping = aes(x = carat, y = price)) +
    geom_point()
```

*(Hint: You might need to look at the help for each function to find out the argument names)*

## 4.

**Submit your work** (1 pt)

Make sure you Render your completed file once you've completed all the tasks. Double-check the PDF file contains all your solutions, then submit both the Quarto file (`homework_01.qmd`) and the PDF file (`homework_01.pdf`).
